---
title: "Stat 123 Summer 2023 Midterm 1b"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All answers are to be done with R commands.

## Question 1

Download the data file artists.csv from Brightspace and save it in a directory of your choice.

Read the data file as a data frame and name it artist.


```{r q1}

#a. Find the sum of the artists' height and display the result below.
artists <- read.csv('artists.csv')
head(artists)

#b. Find the average of the artists' weight and display the result below.
artists_weight <- artists[, "Weight"]
avg_artsits_weigth <- mean(artists_weight)
avg_artsits_weigth


```

## Question 2

You will use the built-in data chickwts and the dplyr package for this question.

```{r q2}

#a. Load the dplyr package using a R command.

library(dplyr)

#b. Display the first 8 observations (rows) in the dataset.
chick <- chickwts
chickwts[1:8, ]


#. For the following questions, you MUST use dplyr piping commands to complete the tasks.

#c. Find out how many chickens receive the meatmeal feed.
meal <- chickwts['feed'] %>%
  filter(feed == 'meatmeal')

length(meal[, 'feed'])


#d. Create a new data frame from chickwts named chickwt2 that contains chickens with
#   weight more than 150. Do not show the results.
chickwt2 <- chickwts['weight']  %>% filter(weight > 150)
length(chickwt2[,'weight'])


#e. Display the chickwt2 data frame in ascending order by weight.
chickwt2_sorted <- chickwt2 %>%
  arrange(weight)

chickwt2_sorted

#f. Display the mean weight of chickens grouped by their feed type.
group <- chickwts %>%
  group_by(feed) %>%
  summarize(mean_weight = mean(weight))

group

```


Once you are satisfied with your results, knit this file as a html and submit it to Brightspace. If you are running out of time and the file does not knit, add the character # at the beginning of the commands that did not run. Then knit it as a html file and submit.

