---
title: "assignment 3"
author: "Raul Rodriguez"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1


```{r}

df <- read.csv('homework3Data.csv')

# a)
# Load the ggplot2 library
library(ggplot2)



ggplot_factory <- function(df, df_col, title_name) {
  ggplot(data = df, aes(x = .data[[df_col]])) +
    geom_histogram(fill = 'gray', color = 'black') +
    geom_vline(aes(xintercept = mean(.data[[df_col]])), color = 'red') +
    geom_vline(aes(xintercept = median(.data[[df_col]])), color = 'blue') +
    geom_density(color = 'green') +
    labs(title = title_name)
}


a_plot <- ggplot_factory(
  df, 
  df_col = 'A', 
  title_name = 'Column A')

b_plot <- ggplot_factory(
  df, 
  df_col = 'B', 
  title_name = 'Column B')

c_plot <- ggplot_factory(
  df, 
  df_col = 'C', 
  title_name = 'Column C')

c_plot <- ggplot_factory(
  df, 
  df_col = 'C', 
  title_name = 'Column C')

x_plot <- ggplot_factory(
  df, 
  df_col = 'X', 
  title_name = 'Column X')

y_plot <- ggplot_factory(
  df, 
  df_col = 'Y', 
  title_name = 'Column Y')

z_plot <- ggplot_factory(
  df, 
  df_col = 'Z', 
  title_name = 'Column Z')




a_plot
b_plot
c_plot

x_plot
y_plot
z_plot

#b) 
# answered in text 

#c) 

mean_X <- mean(df$X)
sd_X <- sd(df$X)

interval_68 <- c(mean_X - sd_X, mean_X + sd_X)
interval_95 <- c(mean_X - 2 * sd_X, mean_X + 2 * sd_X)
interval_997 <- c(mean_X - 3 * sd_X, mean_X + 3 * sd_X)

interval_68
interval_95
interval_997

# d) 
quantiles <- quantile(df$X, c(0.003, 0.025, 0.1587, 0.5, 0.8413, 0.975, 0.997))
quantiles

# e)
intervals_qnorm <- qnorm(c(0.0015, 0.025, 0.1587, 0.5, 0.8413, 0.975, 0.9985), mean = mean_X, sd = sd_X)

intervals_qnorm


# f)

#pop mean estimate
sample_mean <- mean(df$X)
sample_mean

#sd
sample_sd <- sd(df$X)
sample_size <- length(df$X)
#estimated error
estimated_se <- sample_sd / sqrt(sample_size)
estimated_se
#critical value
diff <- sample_size - 1
critical_value_t <- qt(0.86, diff)
critical_value_t

confidence_interval_lower <- sample_mean - (critical_value_t * estimated_se)
confidence_interval_upper <- sample_mean + (critical_value_t * estimated_se)

confidence_interval_lower
confidence_interval_upper

```
b) Without any formal tests and purely based on visual aid, column x appears to be normally distributed because the graph closely conforms with the normal distribution bell curve.

e) comparing these intervals, we can see that the intervals obtained from part (e) using qnorm() are closest to the intervals from part (c) calculated using the mean and standard deviation. 

## Question 2


```{r}

library(dplyr)
# a) 
names <- starwars$name


# b) 
nchar(names[c(7, 19, 31)])

# c)
num_char <- numeric(length(names))

for (i in 1:length(names)) {
  num_char[i] <- nchar(names[i])
}


# d)
num_char <- sapply(names[c(7, 19, 31)], nchar)



```

## Question 3


```{r}
library(ggplot2)
data <- read.csv("homework3Data.csv")
set.seed(123)

# a)
n_bootstrap <- 10000
sample_size <- 250
n_replicate <- 10000

bootstrap_medians <- numeric(n_bootstrap)

for (i in 1:n_bootstrap) {
  bootstrap_sample <- sample(data$Y, size = sample_size, replace = TRUE)
  bootstrap_medians[i] <- median(bootstrap_sample)
}

# b)
ggplot(data.frame(bootstrap_medians), aes(x = bootstrap_medians)) +
  geom_histogram(fill = "gray", color = "black", bins = 30) +
  labs(x = "Median of Y", y = "Frequency", title = "Sampling Distribution of Median")

# c)
estimated_median <- median(data$Y)

# d)
lower_ci <- quantile(bootstrap_medians, 0.025)
upper_ci <- quantile(bootstrap_medians, 0.975)

```
