---
title: "homework 2"
author: "Raul Rodriguez"
date: "2023-06-21"
output: html_document
---

****---
title: "Lab 5"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1


```{r}
library(dplyr)

# a) 
#dept a 
#grouped by admit and gender
dept_a <-UCBAdmissions[,,'A']
ncol(dept_a)
ncol(dept_a)
class(dept_a)

# b) 
admit_gender <- rowSums(UCBAdmissions, dims = 2)
admit_gender



# c) 

total_applicants <- rowSums(admit_gender)
total_applicants
# d) 
admit_rate <- admit_gender / total_applicants

admit_rate

```
e) After taking a look at the results it can be observed that both the admission
and rejection rate for male and female participants is equivalent. Both have 
proportional ratios to according to the percentage they  corresponding size of the population. 

## Question 2


```{r}
library(ggplot2)

data = read.csv('Government_expenditure_per_student.csv')

#a)
ggplot(data, aes(x = Value)) +
  geom_histogram(fill = "steelblue", color = "white") +
  labs(title = "Distribution of Government Expenditure per Student",
       x = "Value")

#b) answered outside of chunk

#c)

# Compute the center value
mean_value <- mean(data$Value)
median_value <- median(data$Value)

# Compute the measures of variability
sd_value <- sd(data$Value)
range_value <- max(data$Value) - min(data$Value)

mean_value
median_value
sd_value
range_value

#d) 
trimmed_mean <- mean(data$Value, trim = 0.05)
trimmed_mean
```
b) The data distribution shows the data to have a right-skew. 


## Question 3

```{r}
#a)
pop <- c(1:30)
pop_mean <- mean(pop)
pop_sd <- sqrt((30-1)/30) * sd(pop)

pop
pop_mean
pop_sd

#b)
samp <- sample(pop, size = 100, replace = TRUE)

#c)


#x (pop_mean): sample mean

#t: the t-critical value

#s (pop_sd): sample standard deviation

#n (length(pop)): sample size

margin <- qt(0.975,df = length(pop)-1)* pop_sd / sqrt(length(pop))


lowerinterval <- pop_mean - margin
upperinterval <- pop_mean + margin

lowerinterval
upperinterval

#d) answered outside of chunk


#e)

```

d) My confidence interval uses the population and the margin to output the desired confidence levels. But the computed result does not contain the population mean.