---
title: "Lab 8"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## 1. Read in data file and set up data frame for analysis.

```{r}

realestate <- read.csv('RealEstate.csv')
head(realestate)

#1a.
house <- data.frame(
  realestate$X2.house.age,       realestate$X3.distance.to.the.nearest.MRT.station,
realestate$X4.number.of.convenience.stores,
realestate$Y.house.price.of.unit.area)

#1b.
colnames(house) <- c('age','mrt.dist','num.conv.stores','price')

# You can also use dplyr to do this.

```

```{r}
library(ggplot2)
#1c price vs age

plot_age <- ggplot(house, aes(x = age, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(x = "Age", y = "Price", title = "Scatterplot: Age vs. Price")



plot_age

```



```{r}

#1c price vs distance to mrt station

plot_mrt_dist <- ggplot(house, aes(x = mrt.dist, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(x = "Distance to MRT Station", y = "Price", title = "Scatterplot: Distance to MRT vs. Price")

plot_mrt_dist

```


```{r}

#1c price vs no. of convenience stores nearby

plot_num_conv_stores <- ggplot(house, aes(x = num.conv.stores, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(x = "Number of Convenience Stores", y = "Price", title = "Scatterplot: Num. of Conv. Stores vs. Price")

plot_num_conv_stores

```

```{r}

#1d Brief discussion about the plots


```
age vs price: the price seems to reach a 'baseline' between 'young' houses and 'older' houses. But younger houses are slightly more expensive than the baseline while the oldest houses seem to have a dramatic price spike. Meaning there is a link: the newer a house is the pricier it gets and this price diminishes until a certain extent. Once you take a look at really old houses the price spikes again. 

MRT Station vs Price: This positively correlates the distance from the MRT Station with a house's price. Meaning that the closer a house is to the station the more it increases in value.

Number of convenience store to price: There seems to be a linear relation between the number of convenience stores and the price of a house. It seems that the more stores there are, the pricier a house becomes. 

## 2. Individual correlations.

```{r}


cor_age_price <- cor(house$age, house$price)

cor_mrt_dist_price <- cor(house$mrt.dist, house$price)

cor_num_conv_stores_price <- cor(house$num.conv.stores, house$price)

cor_age_price
cor_mrt_dist_price
cor_num_conv_stores_price

```

## 3. Correlation matrix for all variables.

```{r}

cor(house)

# The y = house specification asks R to compute all correlations in the house
# data frame.

house |> cor(y = house, method = "pearson") |>  round(3)

# The correlation coefficent between number of convenience stores and distance
# to mrt station is -0.603.


```
The correlation suggests that there is a negative correlation between the number of convenienve stores and the MRT Station distance. This means that as the MRT distance increases the number of convenience stores decreases.  

