---
title: "STAT 260 R Assignment 2"
author: "Rodriguez Castro Raul V01030827"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{Rodriguez Castro, Raul V01030827}
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---
P(X = 2) -> dbinom(2, size=18, prob=0.171)
P(X ≤ 3) -> pbinom(3, size=18, prob=0.171)

# Question 1
```{r 1k}
lambda = 4.5 * 7.5 #seconds * duration
```
  ## Section a
  P(X <= 35) i.e cdf

```{r 1a}
  
  ppois(q = 35, lambda = lambda)
```
  ## Section b
  
  P(X = 33) i.e pmf (discrete) or pdf (continuous)
```{r 1b}
  dpois(33, lambda)
```
  ## Section c
  P(30 <= X <= 36) = P(X <= 36) - P(X <= 29)
  scale down to sample space
```{r 1c}
ppois(36, lambda) - ppois(29, lambda)

```

# Question 2
```{r 2}
  blades_total = 196
  blades_prob = 0.11
```
  ## Section a
```{r 2a}
  # P(X >= 20) = 1 - P(X <= 19)

  at_least_20 = 1 - pbinom(19, size = blades_total, prob = blades_prob)

  #P(X = 20)
  equals_20 = dbinom(20, size = blades_total, prob = at_least_20)
  
```
  ## Section b
```{r 2b}
x = seq(0,50, by = 1)
y = dbinom(x, size = blades_total, prob = 0.11)

plot(x, y, type = 'h',
     main = 'Title',
     xlab = 'X title'
     )
grid()
```
  ## Section c
```{r 2c}
  #(P X >= 20) = 1 - P(X >= 19)
  
  s = sqrt(blades_total * blades_prob * (1 - blades_prob)) # sqrt(npq)
  mu = blades_total * blades_prob

  approximation = 1 - pnorm(19, mean = mu, sd = s)
  approximation
 
  #1 - pbinom(19, size = blades_total, prob = blades_prob)

```
  ## Section d
```{r 2d}
    
    random_X = dnorm(x, mean = mu, sd = s)
    plot(x, random_X, type = 'h',
     main = 'Probability They Get Replaced',
     xlab = 'number of blades')
    
```

# Question 3

  ## Section a
  
  P(23.7 <= X <= 30.4) = P(X <= 30.4) - P(X <= 23.7)
  note self: continuous is inclusive P(X <= 23.7)
```{r 3a}
  mean = 28.3
  sd = 2.39
  x =  23.7
  
  pnorm(30.4, mean = mean, sd = sd) - pnorm(x, mean = mean, sd = sd)
    
    
```
  
  ## Section b
  
  P(X >= 27.4) = 1- P(X <= 27.3)
```{r 3b}
  1 - pnorm(27.4, mean = mean, sd = sd)
```
 
  ## Section c
  
  P(25 <= X <= 31.6) = P(X <= 31.6) - P(X <= 25)
```{r 3c}
  pnorm(31.6, mean = mean, sd = sd) - pnorm(25, mean = mean, sd = sd)
  
```
  ## Section d
```{r 3d}
  qnorm(0.35, mean = mean, sd = sd)
```

# Question 4
  ## Section a
```{r 4a}
  set.seed(111)
  simulation.data = rbinom(2800, size = 72, prob = 0.36)
```
  ## Section b
```{r 4b}
  hist(simulation.data, 
     main = "Histogram of Simulation Data",
     xlab = "Values of X",
     ylab = "Frequency",
     col = "skyblue",  
     breaks = 20) 
```
  Histogram is normally distributed with a slight right skew.
  
  ## Section c
```{r 4c}
  mean(simulation.data)
```
The simulation is very close to the expected value indicating that the sample mean is a reliable estimate of the population mean.
