---
title: "STAT 260 R Assignment 3"
author: "Rodriguez Castro Raul V01030827"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{Rodriguez Castro, Raul V01030827}
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

# Question 1 

```{r 1a}
  repair_time <- c(159, 224, 222, 149, 280, 379, 362, 260, 101, 179, 168, 485, 212, 264, 250, 170)
  sample_mean <- mean(repair_time)
  sample_sd <- sd(repair_time)
  t_score <- qt(0.975, df = length(repair_time) - 1)
  margin_of_error <- t_score * (sample_sd / sqrt(length(repair_time)))
  
  lower_bound <- sample_mean - margin_of_error
  upper_bound <- sample_mean + margin_of_error
  
  lower_bound
  upper_bound
```

```{r 1b}
  t_test_result <- t.test(repair_time, mu = 225, alternative = "greater")
  t_test_result
```

c) t = 0.66852

d) p-value = 0.257

e) we fail to reject the null hypothesis since there is not enough evidence.

f) (a) and (e) rely on the assumption that the repair times follow a normal distribution.


# Question 2

```{r 2a}

  meanX=NULL
  for (i in 1:1000){
    X<-rgamma(n=200,shape=3, scale=2)
    meanX[i]=mean(X)
  }

  hist(meanX, main = "meanX", xlab = "Mean Value", ylab = "Frequency")
```
a) The plot looks approximately normally distributed with a small left tail and a slightly bigger right tail.

```{r 2b}
  meanX <- NULL


for (i in 1:1000) {
  X <- runif(n = 200, min = 0, max = 1)
  meanX[i] <- mean(X)
}
hist(meanX, main = "meanX", xlab = "Mean Value", ylab = "Frequency")
```
c) The plot looks approximately normally distributed with a small left tail and a slightly bigger right tail.

# Question 3

```{r 3a}
F1 = c(177, 176, 198, 197, 185, 188, 206)
F2 = c(206, 193, 192, 188, 207, 210, 197)

sd_F1 <- sd(F1)
sd_F2 <- sd(F2)

threshold <- 0.1 * max(sd_F1, sd_F2)
is_ratio_close <- (abs(sd_F1 - sd_F2) <= threshold)
is_ratio_close
sd_F1
sd_F2
```

a) Since the standard deviations are not close to each other it is more appropriate to use unpooled procedures.

``` {r 3b}
t_test_result <- t.test(F1, F2, alternative = "less", var.equal = FALSE)
t_test_result
```

c) p-value = 0.05283

d) There is not enough evidence to reject the null hypothesis.